---
title: "ETM538 HW 4"
author: "Jordan Hilton"
date: "February 3, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initial Setup

Here's a modified version of the provided data load:

```{r dataload}

mem_claims             <- read.csv("Claims_Y1.csv")
mem_days               <- read.csv("DayInHospital_Y2.csv")
mem_info               <- read.csv("Members_Y1.csv")
risk_model             <- read.csv("risk_model_1.csv")
```


Here's the processing code:

```{r processing}
colnames(mem_days) <- c("MemberID", "Days")
mem_to_risk <- merge(mem_days, risk_model, by="Days")
claims_to_risk <- merge(mem_claims, mem_to_risk, by = "MemberID")
```

Here's the code to calculate the a priori probabilities:

```{r}
n_claims <- length(claims_to_risk[,1])     # note that we have to pick a column.

risks <- as.data.frame(as.character(claims_to_risk$RiskLevel))

riskl <- as.list(risks)

risk_count <- aggregate(risks, riskl, FUN=length)

colnames(risk_count) <- c("RiskLevel", "RiskCount")

a_priori <- risk_count

a_priori$Total <- n_claims

a_priori$Prob <- a_priori$RiskCount / n_claims

colnames(a_priori) <- c("RiskLevel", "RiskCount","Total", "Prob") ##Modified this from original code

write.csv(a_priori, file = "out_a_priori.csv", row.names = FALSE)
```

The calculation for the condition on Charlson index:

```{r charlsonindex}
on_charlson <- data.frame(as.character(claims_to_risk$RiskLevel),
                             as.character(claims_to_risk$CharlsonIndex))

colnames(on_charlson) <- c("RiskLevel", "Charlson")

df_char <- as.data.frame(as.character(on_charlson$Charlson))

colnames(df_char) <- c("Charlson")

l_char <- on_charlson$Charlson

l_risk <- on_charlson$RiskLevel

count_char <- aggregate(df_char, by=list(l_char, l_risk), FUN=length)

colnames(count_char) <- c("Charlson", "RiskLevel", "Count")

# check the total to make sure everything is present and accounted for.

n_chars <- sum(count_char$Count)

n_missing <- n_claims - n_chars

print(paste("A Posteriori Charlson -- ", toString(n_missing), " are missing."))

post_char <- merge(count_char, risk_count, by="RiskLevel")

post_char$Prob <- post_char$Count / post_char$RiskCount

post_char$Label <- paste(post_char$Charlson, post_char$RiskLevel, sep="|")

# reorder the columns

post_char <- post_char[c("Label", "Charlson", "RiskLevel", "Count", "RiskCount", "Prob")]
```

The calculation for the condition on length of (first) stay:

```{r length}
# extract length of stay as a vector

new_stay <- as.character(claims_to_risk$LengthOfStay)

# assign default value to missing columns

new_stay[new_stay == ''] <- '0 days'

on_stay <- data.frame(as.character(claims_to_risk$RiskLevel),
                          as.character(new_stay))

colnames(on_stay) <- c("RiskLevel", "Stay")

df_stay <- as.data.frame(as.character(on_stay$Stay))

colnames(df_stay) <- c("Stay")

l_stay <- on_stay$Stay

l_risk <- on_stay$RiskLevel

count_stay <- aggregate(df_stay, by=list(l_stay, l_risk), FUN=length)

colnames(count_stay) <- c("Stay", "RiskLevel", "Count")

# check the total to make sure everything is present and accounted for.

n_stays <- sum(count_stay$Count)

n_missing <- n_claims - n_stays

print(paste("A Posteriori stay -- ", toString(n_missing), " are missing."))

post_stay <- merge(count_stay, risk_count, by="RiskLevel")

post_stay$Prob <- post_stay$Count / post_stay$RiskCount

post_stay$Label <- paste(post_stay$Stay, post_stay$RiskLevel, sep="|")

# reorder the columns

post_stay <- post_stay[c("Label", "Stay", "RiskLevel", "Count", "RiskCount", "Prob")]
```

The calculation for the condition on primary condition group:

```{r primaryconditiongroup}

on_pcg <- data.frame(as.character(claims_to_risk$RiskLevel),
                      as.character(claims_to_risk$PrimaryConditionGroup))

colnames(on_pcg) <- c("RiskLevel", "pcg")

df_pcg <- as.data.frame(as.character(on_pcg$pcg))

colnames(df_pcg) <- c("pcg")

l_pcg <- on_pcg$pcg

l_risk <- on_pcg$RiskLevel

count_pcg <- aggregate(df_pcg, by=list(l_pcg, l_risk), FUN=length)

colnames(count_pcg) <- c("pcg", "RiskLevel", "Count")

# check the total to make sure everything is present and accounted for.

n_pcgs <- sum(count_pcg$Count)

n_missing <- n_claims - n_pcgs

print(paste("A Posteriori pcg -- ", toString(n_missing), " are missing."))

post_pcg <- merge(count_pcg, risk_count, by="RiskLevel")

post_pcg$Prob <- post_pcg$Count / post_pcg$RiskCount

post_pcg$Label <- paste(post_pcg$pcg, post_pcg$RiskLevel, sep="|")

# reorder the columns

post_pcg <- post_pcg[c("Label", "pcg", "RiskLevel", "Count", "RiskCount", "Prob")]
```

Writing out to csvs:

```{r write}
write.csv(post_char, file="out_charlson.csv")
write.csv(post_stay, file="out_stay.csv")
write.csv(post_pcg, file="out_pcg.csv")
```



## Question responses:

### 1. Dataset explanation
 
Explain how the R data pipeline works, by describing the role of each of the following R data sets:

* mem_to_risk - Combines the "Days in Hospital Y2" and "Risk Level" tables to provide information about the risk level of each member ID.
* claims_to_risk - Combines the "Claims Y1" table with the mem-to-risk table, so that now we have a single table with information about the claims in Y1 and the risk level in Y2 for each member.
* risk_count - A simple count of how many rows there are in claims-to-risk for each different risk level.
* a_priori - Contains the a priori probability for each risk level calculated from claims-to-risk.
+ on_charlson - A subset of the claims-to-risk table containing only the Charlson index and risk level fields
+ count_char - Counts the number of claims for each possible combination of risk level and Charlson index in claims-to-risk.
+ post_char - Gets the total number of risk counts for each risk level, then uses the count from count_char to calculate the a posteriori probability

### 2. Spreadsheet operation

Charlson index seems to be the most influential variable- when you hold other variables constant and change Charlson index, you get the biggest change in predicted outcome.

### 3. Label column

The label column is produced in the R code by this line, which pastes together the risk level and the independent variable we're looking at:

```{r labelexample}
post_pcg$Label <- paste(post_pcg$pcg, post_pcg$RiskLevel, sep="|")
```

It's used by the Excel sheet as the key for the vlookup function to reference the chosen combination of independent variable and risk level for the naive Bayes calculation.

### 4. Missing values for long hospital stays

From the lecture, the way we shoudl do this is to add 1 count for every possible combination of independent variable and risk level to our probably calculator.

### 5. Unlikely B

First, here are some example combinations that predict risk levels A, C, and D:

* A
* C 
* D Charlson:3-4	Stay:4- 8 weeks	PCG:SEPSIS predicts for D at 67.2%

Looking at the probability tables, what's happening is just that because that the counts for risk level B are lower than for other categories- presumably because the odds of spending exactly 2 days in the hospital are lower than the odds of spending either 1 day or between 3 and 5 days.

### 6. Changed Risk buckets

### 7. Adding a variable
