---
title: "Untitled"
author: "Mengyu Li"
date: "January 26, 2019"
output: html_document
---

## Load packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, error =F)
list.of.packages <- c("dplyr", "rlang", "reshape2","broom", "knitr", "readxl", "tidyverse")
lapply(list.of.packages,function(x){library(x,character.only=TRUE)})
```


## Read raw data from csv
```{r}
df <- read_csv("Heritage.csv")
str(df)
rawData <- df
```
## Prepare data
```{r}
df <- rawData
# select independent and dependent variables
iv_fac <- c('specialty', 'placesvc', 'LengthOfStay', 'dsfs', 'PrimaryConditionGroup', 'CharlsonIndex','sex', 'AgeAtFirstClaim') 
iv_num <- c('paydelay')
dv <- c('DaysInHospital_Y2')
iv <- c(iv_fac, iv_num)
var <- c(iv, dv)
df <- df %>% 
  select(var)
df[is.na(df)] <- 0 # fill NA values by 0

df[iv_fac] <- lapply(df[iv_fac], factor)
str(df)
dfData <- df
```
## Field summary
```{r}
summary(dfData)
```

## A function to return average values for each row in a group set of any combination of variables.
```{r}

fnAvgByGrp <- function(df, dfGrpVar, valVar, rng){
  dfResult <- NULL
  #colName <- c(paste0('X', seq(1:ncol(dfGrpVar ))), paste0('Avg', varVal), 'Condition')
  for (rIdx in 1 : nrow(dfGrpVar)){
    grpVar <- as.character(unlist(dfGrpVar[rIdx,]))
    grpVar
    dfT <- df %>% 
      group_by_at(grpVar) %>%
      summarise(AvgY2 = mean(!! sym(valVar)), nCount = n()) %>% 
      filter(between(AvgY2, rng[1], rng[2])) %>% 
      arrange(desc(AvgY2)) %>% 
      mutate(CONDITION = paste0(grpVar, collapse = ' AND '))
    if (nrow(dfT) > 0){
      dfResult <- bind_rows(dfResult, dfT)
    }
  }
  return(dfResult)
}
```

## Question 2: What is the best 1-condition rule you can find?  
```{r}
df <- dfData
iv # independent variable(s)
nCondition <- 1 # number of condition in rule
dfGrpVar <- data.frame(t(combn(c(iv), nCondition))) # group vraible(s)
valVar <- c('DaysInHospital_Y2') # Value variable
rng <- c(4, 500) # range: lower limit and upper limit

res <- fnAvgByGrp(df, dfGrpVar, valVar, rng)

if (! is.null(res)) {
  res <- arrange(res, desc(AvgY2), desc(nCount))
}
res
res1 <- res
```

## Question 3(1): What is the best 2-condition rule you can find?  
```{r}
df <- dfData
iv
nCondition <- 2
dfGrpVar <- data.frame(t(combn(c(iv), nCondition)))
valVar <- c('DaysInHospital_Y2')
rng <- c(6, 500)

res <- fnAvgByGrp(df, dfGrpVar, valVar, rng)
res <- arrange(res, desc(AvgY2), desc(nCount))# arrange based on the largest AvgY2 & nCount
res
res2 <- res
```

## Question 3(2): What is the best 3-condition rule you can find?  
```{r}
df <- dfData
iv
nCondition <- 3
dfGrpVar <- data.frame(t(combn(c(iv), nCondition)))
valVar <- c('DaysInHospital_Y2')
rng <- c(6, 500)

res <- fnAvgByGrp(df, dfGrpVar, valVar, rng)
res <- arrange(res, desc(AvgY2), desc(nCount))
res
res3 <- res
```
## 4.	Can you find 12 different rules that predict high risk situations as defined above?  No rule is allowed to have more than 3 conditions.  No more than 3 of the rules can use exactly the same set of fields.  For instance, if you have 5 rules that predict high risk situations using the fields dsfs, charlson index, and paydelay, you can only use three of them in your list.  However, you can also include 3 more tha use, for example, dsfs, charlson index, and placesvc.  



```{r}
df <- res3
head(df)

df %>% 
  group_by(CONDITION) %>% 
  mutate(rank = rank(desc(AvgY2), ties.method = 'first')) %>% 
  filter(rank <=3) #No more than 3 of the rules can use exactly the same set of fields.  
df[1:12,] #12 different rules that predict high risk situations 
  

```


