---
title: "Homework 1"
author:
   - Andey Nunes, MS
   - Jordan Hilton
   - additional team member name(s) here
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

## Document Setup
The first step for this week is to set up the R Markdown document options.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <- c("broom", "pander", "readxl", "tidyverse")
lapply(packages, require, character.only = T)

options(digits = 2, scipen = 999, stringsAsFactors = F)
```

Next step, load the data sets for the homework and summarize.

```{r load data}
catalog <- read_excel("catalog.xls")
customers <- read_excel("customers.xls")
order_lines <- read_excel("order_lines.xls")
orders <- read_excel("orders.xls")
```
### Summary tables
```{r summaries}

catalog_summary <- summary(catalog)
glimpse(catalog)
pander(catalog_summary, caption = "catalog summary table")

customers_summary <- summary(customers)
glimpse(customers)
pander(customers_summary, caption = "customers summary table")

order_lines_summary <- summary(order_lines)
glimpse(order_lines)
pander(order_lines_summary, caption = "order_lines summary table")

orders_summary <- summary(orders)
glimpse(orders)
pander(orders_summary, caption = "orders summary table")
```

| column names (variables) | assign a type: “question”, “answer”, or “link”  | variable class | count missing values | range = max - min |

#### This section is for building some custom functions that will come in handy later
```{r custom functions}
countNA <- function(x) { sum(is.na(x)) }
get_range <- function(x) {diff(range(x))}
```



## Homework Questions




## References
