---
html_document:
  df_print: paged
  toc: yes
  toc_depth: '3'
author: "Andey Nunes, Mengyu Li, Jordan Hilton, Peter Boss"
date: "2/2/2019"
title: "Homework 3"
---

#### Document and Exercise Set Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <- c("knitr","pander", "tidyverse")

lapply(packages, require, character.only = T)

options(digits = 2, scipen = 999)
```


```{r load data}
Claims_and_Days <- read_csv("Claims_and_Days.csv",
col_types = cols(CharlsonIndex = col_character()))

names(Claims_and_Days)
```

```{r clean data order factors}
Claims_and_Days <- Claims_and_Days %>%
  mutate(CharlsonIndex = case_when(
    CharlsonIndex == "0" ~ "0" ,
    CharlsonIndex == "2-Jan" ~ "1-2",
    CharlsonIndex == "4-Mar" ~ "3-4",
    CharlsonIndex == "5+" ~ "5+"),
    AgeAtFirstClaim = case_when(
      AgeAtFirstClaim == "19-Oct" ~ "10-19",
      AgeAtFirstClaim != "19-Oct" ~ AgeAtFirstClaim),
    LengthOfStay = if_else(is.na(LengthOfStay), "0 or unknown", LengthOfStay)) %>%
  mutate(CharlsonIndex = ordered(CharlsonIndex, levels = 
                                   c("0","1-2", "3-4", "5+")),
         AgeAtFirstClaim = ordered(AgeAtFirstClaim, levels = c(
           "0-9", "10-19", "20-29", "30-39", "40-49", "50-59", "60-69", "70-79",  "80+")),
         LengthOfStay = ordered(LengthOfStay, levels = c(
           "0 or unknown", "1 day", "2 days", "3 days","4 days", "5 days", "6 days", "1- 2 weeks", "2- 4 weeks", "4- 8 weeks", "8-12 weeks", "12-26 weeks", "26+ weeks"))
)
```


# Tidy 1 R Algorithm: Homework 3

## Quantize the answer field: `Risk_Level`

```{r risk level Vlookup tab}
VlookupSim <- Claims_and_Days %>%
  mutate(Risk_Level = case_when(
    DaysInHospital_Y2 == 0 ~ "A",
    DaysInHospital_Y2 == 1 ~ "B",
    DaysInHospital_Y2 == 2 | DaysInHospital_Y2 == 3 ~ "C",
    DaysInHospital_Y2 == 4 | DaysInHospital_Y2 == 5 ~ "D",
    DaysInHospital_Y2 >= 6 ~ "E"),
      Risk_Level_label = case_when(
        Risk_Level == "A" ~ "no risk",
        Risk_Level == "B" ~ "very low risk",
        Risk_Level == "C" ~ "low risk",
        Risk_Level == "D" ~ "medium risk", 
        Risk_Level == "E" ~ "high risk")) %>%
  mutate(Risk_Level_label = ordered(Risk_Level_label, levels = c(
    "no risk", "very low risk", "low risk", "medium risk", "high risk")))

str(VlookupSim)
```


## Create pivot tables 

**proportion of observations for risk groups**
```{r prop_obs table function}
prop_obs <- function(df = VlookupSim, x, 
                     y = VlookupSim$Risk_Level_label) {
   prop <- with(df, table(x, y))
   prop <- cbind(prop, total = rowSums(prop))
   high_prop <- prop[,5]/prop[,6]*100
   prop <- cbind(prop, high_risk_proportion = high_prop)
   #proptibble <- tibble(prop) %>% 
    #  arrange(desc(high_risk_proportion))
   
   return(prop)
}

```

#### Primary Condition Group Counts and Proportion of High Risk
```{r PrimaryConditionGroup pivot}
pcg <- prop_obs(x = VlookupSim$PrimaryConditionGroup)
kable(pcg)
```
  

#### Charlson Index Group Counts and Proportion of High Risk

```{r Charlson Index Group pivot}
ci <- prop_obs(x = VlookupSim$CharlsonIndex)
kable(ci)
```
  

#### Length of Stay Counts and Proportion of High Risk

```{r Length of Stay pivot, eval=F}
los <- prop_obs(x = VlookupSim$LengthOfStay)
kable(los)
```


## Simulate the 1R Algorithm

```{r ignore risk level A}
data1R <- filter(VlookupSim, Risk_Level != "A")
```


**Ignoring Risk Level A** build the 1R Rule for each of

+ "PrimaryConditionGroup"
+ "CharlsonIndex" 
+ "LengthOfStay"

> Select the Best Rule and Paste in here

## Questions:

### Q4.1
 Why does our selected rule work better (reference error rate)
 
### Q4.2
For the 1R Rule is it better to have a lower or higher cardinality?

### Q4.3
We are ignoring risk level A because it is a subset of the data that is independent of the group we are trying to predict and it consists of nearly 77% of the original observations. By working with a smaller subset that contains only the observations with any risk level other than "no risk", we have a smaller search space for our algorithm.

## Extra Credit

**Ignoring Risk Level A** rebin the risk level buckets and rebuild the 1R Rule and see if there is a difference.



# Appendix

## Risk Group Counts
**graphical inspection function**
```{r ggplot risk_group function}
risk_group_counts <- ggplot(VlookupSim, aes(x = Risk_Level_label)) +
   geom_text(aes(label = ..count..), stat = "count", vjust = -0.25) +
   geom_bar()
```


```{r Primary Condition Group plot,  eval=F}
risk_group_counts +
  facet_wrap(~PrimaryConditionGroup)
```


```{r Charlson Index plot, eval=F}
risk_group_counts +
  facet_wrap(~CharlsonIndex)
```


```{r Length Of Stay plot, eval=F}
risk_group_counts +
  facet_wrap(~LengthOfStay)
```



