---
title: "HW5 NN Notebook"
author:
- Jordan Hilton
- Mengyu Li  
- Peter Boss
- Andey Nunes, MS
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <- c("DiagrammeR", "hclust", "knitr", "sigmoid", "tidyverse")

lapply(packages, require, character.only = T)

options(digits = 2, scipen = 999)
```

# This assignment makes use of the following packages AND dependencies: 

+ `DiagrammeR`
+ 
+ `hclust`
+ `knitr`
+ `sigmoid`
+ `tidyverse`

Prior to rendering this document please update R, RStudio, and the associated packages to ensure optimal experience.

# 5-1 – Neural Networks

```{r, eval=F, include=F}
# we don't need to include this, it needs work
grViz("
digraph class_example {

   # a 'graph' statement
   graph [overlap = true, fontsize = 10]

   node [shape = circle,
        fixedsize = true,
        width = 0.9] // sets as circles
   o_1; o_2; h_0_1; h_1; h_2; x_0_1; x_1; x_2

   # several 'edge' statements
   o_1->h_0_1 [arrowhead = none]
   o_1->h_1 [arrowhead = none]
   o_1->h_2 [arrowhead = none]
   o_2->h_0_1 [arrowhead = none]
   o_2->h_1 [arrowhead = none]
   o_2->h_2 [arrowhead = none]
   h_1->x_0_1 [arrowhead = none]
   h_1->x_1 [arrowhead = none]
   h_1->x_2 [arrowhead = none]
   h_2->x_0_1 [arrowhead = none]
   h_2->x_1 [arrowhead = none]
   h_2->x_2 [arrowhead = none]
   

}
")

```


## Neural Networks part 5-1-1: Forward-propagation

This assignment is simply to do the next epoch of the in-class exercise.

What are the activations of the hidden and output units after forward-propagation of the input x2 = (0, 1)?

```{r in_class exercise info}
# create input as a column vector
Xix2 <- c(1,0,1)

# create weights using row vectors
wjixh1 <- c(-0.4, 0.2, 0.1)
wjixh2 <- c(-0.2, 0.4, -0.1)
wkjho1 <- c(0.1, -0.2, 0.1)
wkjho2 <- c(0.4, -0.1, 0.1)

# activation function from prior exercise
a <- function(x) if (x > 0) {1} else {0}
```

```{r h_layer}
h1 <- sum(Xix2 * wjixh1)
sigh1 <- sigmoid(h1, method = "logistic")
# outh1 <- 1/(1 + log(-h1))
h2 <- sum(Xix2 * wjixh2)
sigh2 <- sigmoid(h2, method = "logistic")
# outh2 <- 1/(1 + log(-h2))
h <- c(1,sigh1,sigh2)
h
```

The hidden layer has activation of each node, here the 1 is the bias, and the other nodes are greater than 0.

```{r output layer}
o1 <- sum(sigh1 * wkjho1)
sigo1 <- sigmoid(o1, method = "logistic")
# outo1 <- 1/(1 + log(-o1))
o2 <- sum(sigh2 * wkjho2)
sigo2 <- sigmoid(o2, method = "logistic")
# outo2 <- 1/(1 + log(-o2))
output <- c(sigo1, sigo2)
output
```



## Neural Networks part 5-1-2: Back-propagation

Suppose x2 has target output t = (.1, .9). What are the weights after back-propagation on this example? Again, use learning rate η = 0.1 and momentum α = 0.9.

```{r back prop on x2}
# create target as a column vector
t <- c(0.1, 0.9)
eta <- 0.1
alpha <- 0.9
# create input as a column vector
Xix2 <- c(1,0,1)

# create weights using row vectors
wjixh1 <- c(-0.4, 0.2, 0.1)
wjixh2 <- c(-0.2, 0.4, -0.1)
wkjho1 <- c(0.1, -0.2, 0.1)
wkjho2 <- c(0.4, -0.1, 0.1)


```


```{r from target output}
# get the node delta using the target and the output nodes
dr_o <- -(t - output)*output*(1 - output)

# get weight deltas using these node deltas and the learning rate
backpass_wkjho1 <- wkjho1 - eta*dr_o[1]
backpass_wkjho2 <- wkjho2 - eta*dr_o[2]
backpass_h1 <- backpass_wkjho1
```









## Neural Networks part 5-1-3: two epochs

For both the initial input (1, 0) and the 2nd input (0, 1), calculate the outputs using both the initial weights, the weights after 1 epoch, and the final weights after 2 epochs. Comment on the results compared to the targets.

#### Epoch 1
```{r full two input exercise info}
# create input as a column vector
Xix1 <- c(1,1,0)

# create weights using row vectors
e1wjixh1 <- c(-0.4, 0.2, 0.1)
e1wjixh2 <- c(-0.2, 0.4, -0.1)
e1wkjho1 <- c(0.1, -0.2, 0.1)
e1wkjho2 <- c(0.4, -0.1, 0.1)


```


```{r epoch 1 h_layer}
e1h1 <- sum(Xix1 * e1wjixh1)
e1sigh1 <- sigmoid(e1h1, method = "logistic")
# outh1 <- 1/(1 + log(-h1))
e1h2 <- sum(Xix1 * e1wjixh2)
e1sigh2 <- sigmoid(e1h2, method = "logistic")
# outh2 <- 1/(1 + log(-h2))
e1h <- c(1,e1sigh1,e1sigh2)
e1h
```


```{r epoch 1 output layer}
e1o1 <- sum(e1sigh1 * e1wkjho1)
e1sigo1 <- sigmoid(e1o1, method = "logistic")
# outo1 <- 1/(1 + log(-o1))
e1o2 <- sum(e1sigh2 * e1wkjho2)
e1sigo2 <- sigmoid(e1o2, method = "logistic")
# outo2 <- 1/(1 + log(-o2))
e1output <- c(e1sigo1, e1sigo2)
e1output
```

#### Epoch 2
```{r epoch 2 input exercise info}
# create input as a column vector
Xix2 <- c(1,0,1)

# create weights using row vectors
e2wjixh1 <- c(-0.4, 0.2, 0.1)
e2wjixh2 <- c(-0.2, 0.4, -0.1)
e2wkjho1 <- c(0.1, -0.2, 0.1)
e2wkjho2 <- c(0.4, -0.1, 0.1)


```


```{r epoch 2 h_layer}
e2h1 <- sum(Xix2 * e2wjixh1)
e2sigh1 <- sigmoid(e2h1, method = "logistic")
# outh1 <- 1/(1 + log(-h1))
e2h2 <- sum(Xix2 * e2wjixh2)
e2sigh2 <- sigmoid(e2h2, method = "logistic")
# outh2 <- 1/(1 + log(-h2))
e2h <- c(1,e2sigh1,e2sigh2)
e2h
```


```{r epoch 2 output layer}
e2o1 <- sum(e2sigh1 * e2wkjho1)
e2sigo1 <- sigmoid(e2o1, method = "logistic")
# outo1 <- 1/(1 + log(-o1))
e2o2 <- sum(e2sigh2 * e2wkjho2)
e2sigo2 <- sigmoid(e2o2, method = "logistic")
# outo2 <- 1/(1 + log(-o2))
e2output <- c(e2sigo1, e2sigo2)
e2output
```

#### Error rates 

We can calculate the total error at the end of each epoch using the squared error function.

> $E_{total} = \sum \frac{1}{2}(target - output)^2$

```{r epoch 1 error}
E1 <- sum(0.5*(t - e1output)^2)
```

The error rate for the first epoch output is `r E1`.

```{r epoch 2 error}
E2 <- sum(0.5*(t - e2output)^2)
```

The error rate for the second epoch output is `r E2`.

The total error rate is `r E1 + E2`.

> Please provide all the answers in the template “ETM 538 Template Winter 2019 Assignment 5 Neural Network”. Provide also the excel table or R code or… that allowed you to fill the template.



# 5-2 – Instance Based Learning

## Part A


#### Review the Instance Based Classification v2 table. This is a slightly improved version of the table we looked at in class. F1 to I1 represent the weights of each attribute. (Here, they all have equal weight). Answer the following questions:


#### What is the total number of possible weather cases that can be submitted for consideration?


#### Of these, how many get a unique answer under the current set of weights? How many get an unambiguous answer?


#### Of the cases that do not get an unambiguous, unique answer, what is the maximum number of answers proposed? How many cases give the maximum?


#### Do the distance calculations in this table represent Euclidean distance, Manhattan distance, both or neither? Explain.


#### What is the function of the weights in cells F1, G1, H1, and I1? What happens if we vary these weights?


#### Can you assign a set of weights in cells F1, G1, H1, and I1 such that you get a unique answer for every one of the possible weather cases? If so, what are the weights? If not, explain how close you can come.



## Part B


#### Review the Instance Based Classification v2 table. This is a slightly improved version of the table we looked at in class.


#### Build two new tables based on the example. One should use Euclidean distance. The other should use Manhattan distance. Keep the weights in cells F1, G1, H1, and I1 – these simply weight the individual attributes. The distance calculations themselves are embedded in the formulas in the region between J2 and J15.


#### For each of the new tables, re-define the distance calculation for the “Outlook” variable to exploit the fact that rainy, overcast, and sunny can be rank ordered. Assign values of 1, 2, and 3 to these values respectively, and calculate the distance by taking the difference. Note that the minimum distance remains zero, while the maximum distance is now 2.


#### For each of the new tables, change the “Temperature” attribute so that it is an integer between 32 and 100 (representing degrees Fahrenheit). For the training data, use the following values:

+ cool = 50
+ mild = 65
+ hot = 80


#### Scale the distance calculation for Temperature so that it does not swamp the other attributes. Given that our “Outlook” variable will have values between 0 and 2, Temperature should have values between 0 and 2 as well. You can accomplish this in any way that seems reasonable for you, such as using a ratio based on maximum or minimum values, or a cap on the maximum.
